%\RequirePackage{snapshot}
\documentclass[10pt,a4paper]{article}
%\RequirePackage{graphics}
\usepackage[dvips]{graphics}
\RequirePackage{lineno} 
\usepackage[pdftex]{color, graphicx}
\usepackage{amsmath, amsthm}
%\usepackage{amssymb}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{xspace}
\usepackage{cite}
\usepackage{ifthen}

\RequirePackage{lineno}
\setlength{\oddsidemargin}{0cm}
\setlength{\textwidth}{16cm}
\setlength{\topmargin}{-15mm}
\setlength{\textheight}{245mm}
\usepackage[pdfauthor={MT2 Team},bookmarks=true]{hyperref} % bookmarks

%\title{\huge{Charge and flavour ratios in SUSY leptonic final states} \\
\title{\huge{  
System Demonstration of CMS Level-1 Tracking Trigger using Associative Memory Approach} \\
\ \  \\
\large{ }}
\author{names}
\vspace{2cm}
\date{\today}

%========================================================================

%\usepackage{topcapt}
%\input{svjour.cls}
%\input{cms-tdr-note.cls}
%\input{cms-tdr.cls}
%%\usepackage{hepparticles}
%%\usepackage{heppennames2}
\usepackage{ptdr-definitions}
%\usepackage{pennames-ital}

\newcommand{\MT}{\ensuremath{M_{\mathrm{T}}}\xspace}
\newcommand{\MTmax}{\ensuremath{\MT^{\mathrm{max}}}\xspace}
\newcommand{\MTmin}{\ensuremath{\MT^{\mathrm{min}}}\xspace}
\newcommand{\SSee}{\ensuremath{\Pepm\Pepm}\xspace}
\newcommand{\SSem}{\ensuremath{\Pepm\PGmpm}\xspace}
\newcommand{\SSmm}{\ensuremath{\PGmpm\PGmpm}\xspace}
\newcommand{\threel}{3\ensuremath{\ell}\xspace}
\newcommand{\Mll}{\ensuremath{M_{\ell\ell}}\xspace}
\newcommand{\MSFOS}{\ensuremath{M_{\mathrm{SFOS}}}\xspace}
\newcommand{\ptlll}{\ensuremath{\pt(\ell\ell\ell)}\xspace}
\newcommand{\ptlllvec}{\ensuremath{\ptvec(\ell\ell\ell)}\xspace}
\newcommand{\ptveclll}{\ensuremath{\ptvec(\ell\ell\ell)}\xspace}
\newcommand{\DPhilllMET}{\ensuremath{\Delta\phi\left(\ptlllvec,\ptvecmiss\right)}\xspace}
\newcommand{\Mlll}{\ensuremath{M_{\ell\ell\ell}}\xspace}
\newcommand{\Mjj}{\ensuremath{M_{\mathrm{jj}}}\xspace}
\newcommand{\theLumi}{\ensuremath{35.9\fbinv}\xspace}
\newcommand{\WHtoWWW}{\ensuremath{\cmsSymbolFace{WH}\to\cmsSymbolFace{WWW}^{(*)}}\xspace}
\newcommand{\Wmp}{\PWmp}
%\newcommand{\Pemp}{\ensuremath{\cmsSymbolFace{e}^{\mp}}\xspace}
\newcommand{\WWW}{\ensuremath{\Wpm\Wpm\Wmp}\xspace}
\newcommand{\SSWW}{\ensuremath{\Wpm\Wpm}\xspace}%use this to avoid useless space between the Wpm
\newcommand{\ttW}{\ensuremath{\ttbar\Wpm}\xspace}%use this to avoid useless space between the Wpm
\newcommand{\ttZ}{\ensuremath{\ttbar\PZ}\xspace}%use this to avoid useless space between the Wpm
\newcommand{\ttV}{\ensuremath{\ttbar\mathrm{V}}\xspace}%use this to avoid useless space between the Wpm
\newcommand{\ttH}{\ensuremath{\ttbar\mathrm{H}}\xspace}%use this to avoid useless space between the Wpm
\newcommand{\eTL}{\ensuremath{\epsilon_{\mathrm{TL}}}\xspace}
\newcommand{\etl}{\eTL}
\newcommand{\ptl}{\ensuremath{\pt^{\ell}}\xspace}
\newcommand{\ptvecl}{\ensuremath{\ptvec^{\,\ell}}\xspace}
\newcommand{\ptlvec}{\ptvecl}
\newcommand{\ptjet}{\ensuremath{\pt^{\text{jet}}}\xspace}
\newcommand{\ptvecjet}{\ensuremath{\ptvec^{\,\text{jet}}}\xspace}
\newcommand{\ptjetvec}{\ptvecjet}
\newcommand{\ptcor}{\ensuremath{\pt^{\text{corr}}}\xspace}

\begin{document}

\hyphenation{had-ron-i-za-tion}
\hyphenation{cal-or-i-me-ter}
\hyphenation{de-vices}
%\input{newcommand}




\maketitle

\begin{abstract}

It has become evident that the development of a Level 1 tracking trigger system is crucial for the success of the high luminosity LHC (HL-LHC) physics program at the CMS experiment. The high pile-up and high occupancy conditions anticipate an input data rate into the tracking trigger system of ~100 Tbps. This, combined with the 
4 $\mu$s latency required the L1 track finding trigger stage introduce an unprecedented set of challenges requiring the most advanced real time processing technology. A silicon-based Level 1 tracking trigger system has never been realized under these conditions. It is therefore mandatory to demonstrate its feasibility at the system level. Recently, the HEP community has made significant progress in the area of fast track finding using ASICs and FPGAs. This includes a successful vertical-slice demonstration for a CMS tracking trigger at Level 1 for the HL-LHC era, developed by a Fermilab/LPC R\&D collaboration. The demonstration solves the data delivery problematic using a full-mesh communication system architecture, implemented using commerical ATCA standard. As for fast pattern recognition algorithm, the demonstration uses associative memory apporach, that can be implemented on specialized ASICs, or on modern FPGAs when the pattern bank size needed is not very large, followed by precision track fitting implemented on dedicated FPGAs. The demonstration system comprises a full trigger path using a single tracking trigger tower (1/48 of the full detector). It uses simulated high-luminosity data as input to measure trigger latency and efficiency, to study overall system performance, and to identify appropriate solutions to possible bottlenecks.  The demonstration system achieved excellent performance in terms of tracking efficiency and momentum resolution, within a very short latency (2.5 $\mu$s). This latency includes the time needed for data delivery to the trigger towers, for pattern recognition, and for track fitting. This demonstration is proof that fast data delivery, fast pattern recognition, and track fitting can be implemented successfully using the full mesh ATCA architecture and an associative memory approach. While the focus here is on the HL-LHC, the techniques developed are rather generic and could have other applications beyond tracking triggers, within and outside HEP. The details of the demonstration will be described, with highlights on the new techniques developed to address the unique challenges at the chip, board, and system level throughout the entire tracking trigger processing chain.  


\end{abstract}

\newpage
%
\tableofcontents
\linenumbers
\newpage

%%%%%%%% Definitions
%\def\eslash{{\hbox{$E$\kern-0.5em\lower-.05ex\hbox{/}\kern-0.10em}}}
%\def\MET{\mbox{$\eslash_T$}}
%\def\hslash{{\hbox{$H$\kern-0.5em\lower-.05ex\hbox{/}\kern-0.10em}}}
%\def\MHT{\mbox{$\hslash_T$\ }}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Begin text %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\include{Introduction}
\include{System_description}
\include{HardwareSetup}
\include{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\appendix


%\clearpage

%
\bibliography{}
\bibliographystyle{lucas_unsrt_epjc}



%----------------------------------------
% Pour la bibliographie
%----------------------------------------

\newpage
\thispagestyle{empty}

\begin{thebibliography}{99}


\bibitem{bib:Rist-89} M. Dell Orso, L. Ristori - {\it "VLSI Structure For Track Finding"} - Nucl.Instr. and Meth. A278 (1989), 436; L.~Ristori and G.~Punzi, Ann.~Rev.~Nucl.~Part. Sci. 60, 595-614 (2010).

\bibitem{bib:Ade-06} J. Adelman {\it et al.} - {\it "Real time secondary vertexting at CDF"} - Nucl.Instr. and Meth. A569 (2006), 111

\bibitem{bib:Ade-07} J. Adelman {\it et al.} - {\it "The Silicon Vertex Trigger upgrade at CDF"} - Nucl.Instr. and Meth. A572 (2007), 361

\bibitem{bib:FTK-TP}  - {\it "FTK: A Hardware Track Finder for the ATLAS Trigger"} - Technical Proposal (2010)

\bibitem{bib:FTK-TDR}  - {\it "FastTracKer(FTK) Technical Design Report"} - CERN-LHCC-2013-007, ATLAS-TDR-021-2013, 2013




\bibitem{bib:GPU} W. ~Ketchum {\it et al.}  - {\it "
Performance study of GPUs in real-time trigger applications for HEP experiments
"}, in Proceedings of the 2nd International Conference on Technology and Instrumentation in Particle Physics (TIPP 2011), vol. 37, 2012, p. 1965. [Online]. 
Available at: 

\url{http://www.sciencedirect.com/science/article/pii/S1875389212019207}


\bibitem{bib:GPU1} A. ~Gianelle {\it et al.}  - {\it "
Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics Experiments
"}:
\url{http://arxiv.org/abs/1312.0917}



\bibitem{bib:PulsarII} J. Olsen, T.~Liu and Y. Okumura - {\it "
A Full Mesh ATCA-based General Purpose Data Processing Board: Pulsar II
"} - FERMILAB-CONF-13-526-CMS-PPD (Nov. 22, 2013), presented at TWEPP 2013 and submitted to JINST:

\url{http://hep.uchicago.edu/~thliu/PulsarII/PulsarII_twepp_paper.pdf}

\bibitem{bib:PulsarII-results} Y. Okumura, J. Olsen, T.~Liu and H. Ying - {\it "
Prototype Performance studies of a Full Mesh ATCA-based General Purpose Data Processing Board
"} - FERMILAB-CONF-13-527-CMS-PPD (Nov. 22, 2013), presented at IEEE NSS 2013: 

\url{http://hep.uchicago.edu/~thliu/PulsarII/PulsarII_IEEE_NSS_paper.pdf}

\bibitem{bib:PulsarII-weblink} J. Olsen, T.~Liu and Y. Okumura - 
Pulsar IIa prototype work web page:

\url{http://www-ppd.fnal.gov/EEDOffice-w/Projects/atca/}

\bibitem{bib:VIP-11} T.~Liu {\it et al.} - {\it "A New Concept of Vertically Integrated Pattern Recognition Associative Memory (VIPRAM)"} - FERMILAB-CONF-11-709-E, Proceedings of TIPP 2011 conference, Volume 37, 2012, Pages 1973: 

\url{http://www.sciencedirect.com/science/article/pii/S1875389212019165}

or 

\url{http://lss.fnal.gov/archive/2011/conf/fermilab-conf-11-709-e.pdf}

\bibitem{bib:VIP-12} T.~Liu {\it et al.} - {\it "Development of 3D Vertically Integrated Pattern Recognition Associative Memory (VIPRAM)"} - FERMILAB-TM-2493-CMS-E-PPD-TD:

\url{http://lss.fnal.gov/archive/test-tm/2000/fermilab-tm-2493-cms-e-ppd-td.pdf}




%\bibitem{bib:Abb-11} D.~Abbaneo {\it for the CMS collaboration} - {\it "Upgrade of the CMS Tracker with tracking trigger"} - CMS CR -2011/284 (2011)

%\bibitem{bib:Koh-87} T. Kohonen - {\it "Content-Addressable Memories"} - 2nd edition, New York, Springer- Verlag (1987)

\bibitem{bib:Tezzaron} {\it "Tezzaron Semiconductor"}:
\url{http://www.tezzaron.com}


%\bibitem{bib:Pag-06} K. Pagiamtzis, A. Sheikholeslami - {\it "Content-Addressable Memory (CAM) Circuits and Archtectures: A Tutorial and Survey"} - IEEE Journal of Solid-State Circuits, Vol. 41, No. 3 (2006), 712


%\bibitem{bib:XML_wiki}\url{https://twiki.cern.ch/twiki/bin/view/CMSPublic/SWGuideGeomDevelopersGuide}

%\bibitem{bib:XML_oscar}\url{https://twiki.cern.ch/twiki/bin/view/CMSPublic/SWGuideOscarProducer}

%\bibitem{bib:G4SHits_cfi}\url{http://cmssw.cvs.cern.ch/cgi-bin/cmssw.cgi/CMSSW/SimG4Core/Application/python/g4SimHits_cfi.py?view=markup}

\bibitem{bib:Ann-06} A. Annovi {\it et al.} - {\it "A VLSI Processor for Fast Track Finding Based on Content Addressable Memories"} - IEEE Transactions on Nuclear Science, vol. 53, no. 4 (2006), 1

\bibitem{bib:Ann-09} A. Annovi {\it et al.} - {\it "The GigaFitter: A next generation track fitter to enhance online tracking performances at CDF"} - Nuclear Science Symposium Conference Record (NSS/MIC), 2009 IEEE (2009), 1143


\bibitem{bib:AMchip04} A. Annovi {\it et al.}  - {\it Variable Resolution Associative Memory for High Energy Physics
"} - Submitted to the Advancements in Nuclear Instrumentation, Measurement Methods and their Applications (ANIMMA), Ghent Belgium, 6-9 June, 2011.


%\bibitem{bib:Kar-91} V.~Karimaki - {\it "Effective circle fitting for particle trajectories"} - Nucl.Instr. and Meth. A305 (1991), 187

\end{thebibliography}
%----------------------------------------


% For use with the tdr script comment \end{document} 
\end{document}






%\bibitem{Catani:2002ny} 
%  S.~Catani, M.~Fontannaz, J.~P.~Guillet and E.~Pilon,
%  %``Cross-section of isolated prompt photons in hadron hadron collisions,''
%  JHEP {\bf 0205}, 028 (2002)
%  [hep-ph/0204023].
%  %%CITATION = HEP-PH/0204023;%%
%%\cite{Aurenche:2006vj}
%\bibitem{Aurenche:2006vj} 
%  P.~Aurenche, M.~Fontannaz, J.~-P.~Guillet, E.~Pilon and M.~Werlen,
%  %``A New critical study of photon production in hadronic collisions,''
%  Phys.\ Rev.\ D {\bf 73}, 094007 (2006)
%  [hep-ph/0602133].
%  %%CITATION = HEP-PH/0602133;%%
%\bibitem{PFT-11-001} CMS Collaboration,   \textit{"Tau Identification in CMS"}, CMS-PAS-TAU-11-001 (2011). 
%\bibitem{EGM-10-006} CMS Collaboration, \textit{Isolated Photon Reconstruction and Identiﬁcation at $\sqrt{s}$ = 7 TeV}, 
%                     CMS Physics Analysis Summary, CMS PAS EGM-10-006
%\bibitem{PAS-BTV-10-001} CMS Collaboration, CMS-PAS-BTV-10-001.
%
%
