\section{Budget Requests for FY2015}

The fully loaded funding request for FY2015 is \$1294K. This consists of \$716K of labor and \$578K of M\&S including travel to CERN. 
The FY2015 requested funds are for: 

\begin{enumerate}
\item Pulsar II pattern recognition mezzanine (Fermilab engineer-III time, 1FTE): \$270,376.00 fully loaded
	\begin{itemize}	
	\item	Pulsar II pattern recognition mezzanine hardware design, compatible with protoVIPRAM-L1CMS chip
        \item       Pulsar II pattern recognition mezzanine firmware (core engine)
	\end{itemize}

\item The design of 2D protoVIPRAM-L1CMS (Fermilab engineer-III time, 1FTE): \$270,376 fully loaded
	\begin{itemize}
	\item	protoVIPRAM-L1CMS chip design dedicated for CMS L1 track trigger
	\end{itemize}

\item ProtoVIPRAM-L1CMS submission: \$260,000.00 
	\begin{itemize}
	\item	Shared cost of Fermilab dedicated MPW run in CY2015: 40\% of \$450,000 total = \$180,000
        \item       Extra processing cost:   \$80,000
	\end{itemize}	

\item Pulsar II ATCA related Hardware needs: \$178,000.00 
	\begin{itemize}
	\item	One ATCA shelf for the core trigger tower at Fermilab teststand
	\item	One crate (Twelve slots) worth of Pulsar II/RTMs and one HUB board
       \item        Ten Pattern Recongintion Mezzanine card prototypes
	\end{itemize}	

\item Pulsar II firmware work (Northwestern junior engineer time, 1FTE): \$70,000.00 fully loaded
	\begin{itemize}
        \item       Pulsar II firmware and system integration 
        \item       Northwestern University will match \$70,000.00
	\end{itemize}

\item Travel for Fermilab Engineers

\begin{itemize}
	\item	Fermilab engineers:  four trips to CERN (one week per visit). \$14,892.00 including all overheads. 
	\end{itemize}	

\item COLA request by Northwestern 

	\begin{itemize}
        \item 12 months of COLA at \$1K per month for CERN-based postdoc Kevin Sung.  
        \item Kevin Sung built and is responsible for the Northwestern Pulsar II test-stand at CERN.  
	\end{itemize}


\item Funding request for TAMU to work on AM simulation 
	\begin{itemize}
	\item	 A new software professional hire (50\% FTE funding request) : \$45,360 including all overheads
        \item       TAMU will match the cost for the other half.
	\item	Travel to attend critical meetings at CERN. Two trips requested, \$7,446.00 including all overheads
	\end{itemize}	

\item Funding request for Florida to work on AM simulation  (waiting for the actual numbers from Jaco tonight).
	\begin{itemize}
	\item	 A new software professional hire (50\% FTE funding request) : \$60,000 including all overheads
        \item       Florida will match the cost for the other half.
	\item	Travel to attend critical meetings at CERN. Two trips requested, \$7,446.00 including all overheads
	\end{itemize}	


\end{enumerate}


